---
title: "20201021-p8133_probset4_jsg2145"
author: "Jared Garfinkel"
date: "10/21/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

Given: Pr(toxicity at dose level 1) = 0.25

Definition: The MTD means that 2/3 or 2/6 participants experience toxicity.

If the first 3 have no instances of toxicity, then there is no MTD.

If the first 3 have 1 instance of toxicity, then 3 more participants are recruited.

If the first 3 have 2 instances of toxicity, then the MTD is reached.

```{r}
pr_tox1 = dbinom(1, 3, 0.25)
pr_tox1
```

There is a `r pr_tox1` probability that 3 more participants will be recruited.

$Pr(X_2 \ge 2 | X_1 = 1) = Pr(X_1 = 1) \cap Pr(X_2 = 1, 2, or 3)$

```{r}
pr_tox3 = 1 - dbinom(0, 3, 0.25)
pr_tox3

pr_toxc= pr_tox1*pr_tox3
```

There is a `r pr_toxc` probability that the MTD will be reached if there is 1 toxicity response in the first 3.

```{r}
pr_tox2 = dbinom(2, 3, 0.25)
pr_tox2
```

There is a `r pr_tox2` probability that the MTD will be reached in the first 3 participants.

So, there is 1 - (`r pr_toxc` + `r pr_tox2`) probability that the dose level 1 will not reach the MTD. 

1 - (`r pr_toxc` + `r pr_tox2`) = `r 1 - (pr_toxc + pr_tox2)`

