---
title: "20201210-p8133_probset13_jsg2145"
author: "Jared Garfinkel"
date: "12/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 13

## Part a

$X_1$ ~ $N(\mu{t_1,~1})$ $\perp$ $X_2$ ~ $N(\mu{t_2,~1})$

Z = $\omega_1X_1$ + $\omega_2X_2$

$\omega_1^2 + \omega_2^2 = 1$

w, t > 0

$Pr(|Z| > z_{0.025}) ~ N(\mu(\omega_1t_1+\omega_2t_2),~2) > z_{0.025}$

```{r}
errI = 1 - pnorm(0.9875, sd = sqrt(2))
```

The type I error rate is `r round(errI, 4)`.

## Part b

The power is the probability that the null will be rejected given that the null is false.

Z ~ $N(\mu(\omega_1t_1 + \omega_2t_2,~2))$

$\frac{Z - \mu(\omega_1t_1 + \omega_2t_2)}{\sqrt{2}}$ ~ $N(0, 1)$

```{r}
errI2 = (1 - pnorm(0.9875))*sqrt(2)
```

The power is `r round(errI2, 4)` + $\mu(\omega_1t_1 + \omega_2t_2)$.

## Part c

By intuition, $\omega_1 = \omega_2 = \sqrt{0.5}$

## Part d

Given: $\mu$ = 0

Z ~ N(0, 2)

$X_1$ ~ N(0, 1)

$Pr(X_1 > z_{0.005}~or~Z > z_{0.02}) \le 0.025$

$Pr(X_1 > z_{0.005})$

$Pr(Z > z_{0.025})$

```{r, include = FALSE, eval = FALSE}
dnorm(0.98, sd = sqrt(2)) # Z > z_0.025, 0.2218814
qnorm(0.98, sd = sqrt(2)) # 2.90444, most strict
qnorm(0.995) # X_1 > z_0.005, 2.575829
1 - pnorm(2.90444) # so, if x_1 > z_0.005, |Z| > z_0.02?
1 - pnorm(2.575829)

# This should add to < 0.25?
dnorm(0.975, sd = sqrt(2))
qnorm(.975, sd = sqrt(2)) # 2.771808, more strict
qnorm(0.995) # 2.575829
1 - pnorm(2.771808)
1 - pnorm(2.575829)
```

The left-hand equation will be smaller because the alpha value is smaller. Therefore, fewer observations will satisfy the inequality in the left-hand equation compared to the right hand equation.

## Part e

```{r}
errI3 = (1-pnorm(0.99))*sqrt(2)
```

The value of the power will be `r round(errI3, 4)` + $\mu(\omega_1t_1 + \omega_2t_2)$.

A smaller value of alpha may affect the power of the test. However, power and type I error measure different things. 

One tests a null when the null is true and the power tests the null when the alternative is true.

The test still depends on $\mu$, $\omega$, and t.