---
title: "20201108-p8133_probset12_jsg2145"
author: "Jared Garfinkel"
date: "11/8/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 12

Calculate the distribution of B using a randomized PTW design.

The first trial has equal probabilities of each treatment.

## P(tr_0)

P(A) = P(B) = 0.5

## P(tr_1)

However, the probability of the next trial depends on the outcome of the first. So,

$p(tr_1 = A|tr_0 = A_{success}) = 2/3 * 0.2 / 2$

$p(tr_1 = A|tr_0 = A_{failure}) = 1/3 * 0.8 / 2$

$p(tr_1 = A|tr_0 = B_{success}) = 1/3 * 0.8 / 2$

$p(tr_1 = A|tr_0 = B_{failure}) = 2/3 * 0.2 / 2$

$p(tr_1 = A)$ = `r 0.5*(2/3 * 0.2 + 1/3 * 0.8 + 1/3 * 0.8 + 2/3 * 0.2)`

$p(tr_1 = B|tr_0 = A_{success}) = 1/3 * 0.2 / 2$

$p(tr_1 = B|tr_0 = A_{failure}) = 2/3 * 0.8 / 2$

$p(tr_1 = B|tr_0 = B_{success}) = 2/3 * 0.8 / 2$

$p(tr_1 = B|tr_0 = B_{failure}) = 1/3 * 0.2 / 2$

$p(tr_1 = B)$ = `r 0.5*(1/3 * 0.2 + 2/3 * 0.8 + 2/3 * 0.8 + 1/3 * 0.2)`

## P(tr_2)

$p(tr_2 = A|tr_x = (A_{success}, A_{success})) = 0.5 * 2/3 * 0.2 * 0.2 * 3/4$

$p(tr_2 = A|tr_x = (A_{success}, A_{failure})) = 0.5 * 2/3 * 0.2 * 0.8 * 2/4$

$p(tr_2 = A|tr_x = (A_{failure}, A_{success})) = 0.5 * 1/3 * 0.2 * 0.8 * 2/4$

$p(tr_2 = A|tr_x = (A_{failure}, A_{failure})) = 0.5 * 1/3 * 0.8 * 0.8 * 1/4$

$p(tr_2 = A|tr_x = (A_{success}, B_{failure})) = 0.5 * 1/3 * 0.2 * 0.2 * 3/4$

$p(tr_2 = A|tr_x = (A_{success}, B_{success})) = 0.5 * 1/3 * 0.2 * 0.8 * 2/4$

$p(tr_2 = A|tr_x = (A_{failure}, B_{failure})) = 0.5 * 2/3 * 0.8 * 0.2 * 2/4$

$p(tr_2 = A|tr_x = (A_{failure}, B_{success})) = 0.5 * 2/3 * 0.8 * 0.8 * 1/4$

$p(tr_2 = A|tr_x = (B_{success}, A_{success})) = 0.5 * 1/3 * 0.8 * 0.2 * 2/4$

$p(tr_2 = A|tr_x = (B_{success}, A_{failure})) = 0.5 * 1/3 * 0.8 * 0.8 * 1/4$

$p(tr_2 = A|tr_x = (B_{failure}, A_{success})) = 0.5 * 2/3 * 0.2 * 0.2 * 3/4$

$p(tr_2 = A|tr_x = (B_{failure}, A_{failure})) = 0.5 * 2/3 * 0.2 * 0.8 * 2/4$

$p(tr_2 = A|tr_x = (B_{success}, B_{failure})) = 0.5 * 2/3 * 0.8 * 0.2 * 2/4$

$p(tr_2 = A|tr_x = (B_{success}, B_{success})) = 0.5 * 2/3 * 0.8 * 0.8 * 1/4$

$p(tr_2 = A|tr_x = (B_{failure}, B_{failure})) = 0.5 * 1/3 * 0.2 * 0.2 * 3/4$

$p(tr_2 = A|tr_x = (B_{failure}, B_{success})) = 0.5 * 1/3 * 0.2 * 0.8 * 2/4$

$p(tr_2 = A)$ = `r 0.5 * 2/3 * 0.2 * 0.2 * 3/4 + 0.5 * 2/3 * 0.2 * 0.8 * 2/4 + 0.5 * 1/3 * 0.2 * 0.8 * 2/4 + 0.5 * 1/3 * 0.8 * 0.8 * 1/4 + 0.5 * 1/3 * 0.2 * 0.2 * 3/4 + 0.5 * 1/3 * 0.2 * 0.8 * 2/4 + 0.5 * 2/3 * 0.8 * 0.2 * 2/4 + 0.5 * 2/3 * 0.8 * 0.8 * 1/4 + 0.5 * 1/3 * 0.8 * 0.2 * 2/4 + 0.5 * 1/3 * 0.8 * 0.8 * 1/4 + 0.5 * 2/3 * 0.2 * 0.2 * 3/4 + 0.5 * 2/3 * 0.2 * 0.8 * 2/4 + 0.5 * 2/3 * 0.8 * 0.2 * 2/4 + 0.5 * 2/3 * 0.8 * 0.8 * 1/4 + 0.5 * 1/3 * 0.2 * 0.2 * 3/4 + 0.5 * 1/3 * 0.2 * 0.8 * 2/4`

$p(tr_2 = B|tr_x = (A_{success}, A_{success})) = 0.5 * 2/3 * 0.2 * 0.2 * 1/4$

$p(tr_2 = B|tr_x = (A_{success}, A_{failure})) = 0.5 * 2/3 * 0.2 * 0.8 * 2/4$

$p(tr_2 = B|tr_x = (A_{failure}, A_{success})) = 0.5 * 1/3 * 0.2 * 0.8 * 2/4$

$p(tr_2 = B|tr_x = (A_{failure}, A_{failure})) = 0.5 * 1/3 * 0.8 * 0.8 * 3/4$

$p(tr_2 = B|tr_x = (A_{success}, B_{failure})) = 0.5 * 1/3 * 0.2 * 0.2 * 1/4$

$p(tr_2 = B|tr_x = (A_{success}, B_{success})) = 0.5 * 1/3 * 0.2 * 0.8 * 2/4$

$p(tr_2 = B|tr_x = (A_{failure}, B_{failure})) = 0.5 * 2/3 * 0.8 * 0.2 * 2/4$

$p(tr_2 = B|tr_x = (A_{failure}, B_{success})) = 0.5 * 2/3 * 0.8 * 0.8 * 3/4$

$p(tr_2 = B|tr_x = (B_{success}, A_{success})) = 0.5 * 1/3 * 0.8 * 0.2 * 2/4$

$p(tr_2 = B|tr_x = (B_{success}, A_{failure})) = 0.5 * 1/3 * 0.8 * 0.8 * 3/4$

$p(tr_2 = B|tr_x = (B_{failure}, A_{success})) = 0.5 * 2/3 * 0.2 * 0.2 * 1/4$

$p(tr_2 = B|tr_x = (B_{failure}, A_{failure})) = 0.5 * 2/3 * 0.2 * 0.8 * 2/4$

$p(tr_2 = B|tr_x = (B_{success}, B_{failure})) = 0.5 * 2/3 * 0.8 * 0.2 * 2/4$

$p(tr_2 = B|tr_x = (B_{success}, B_{success})) = 0.5 * 2/3 * 0.8 * 0.8 * 3/4$

$p(tr_2 = B|tr_x = (B_{failure}, B_{failure})) = 0.5 * 1/3 * 0.2 * 0.2 * 1/4$

$p(tr_2 = B|tr_x = (B_{failure}, B_{success})) = 0.5 * 1/3 * 0.2 * 0.8 * 2/4$

$p(tr_2 = B)$ = `r 0.5 * 2/3 * 0.2 * 0.2 * 1/4 + 0.5 * 2/3 * 0.2 * 0.8 * 2/4 + 0.5 * 1/3 * 0.2 * 0.8 * 2/4 + 0.5 * 1/3 * 0.8 * 0.8 * 3/4 + 0.5 * 1/3 * 0.2 * 0.2 * 1/4 + 0.5 * 1/3 * 0.2 * 0.8 * 2/4 + 0.5 * 2/3 * 0.8 * 0.2 * 2/4 + 0.5 * 2/3 * 0.8 * 0.8 * 3/4 + 0.5 * 1/3 * 0.8 * 0.2 * 2/4 + 0.5 * 1/3 * 0.8 * 0.8 * 3/4 + 0.5 * 2/3 * 0.2 * 0.2 * 1/4 + 0.5 * 2/3 * 0.2 * 0.8 * 2/4 + 0.5 * 2/3 * 0.8 * 0.2 * 2/4 + 0.5 * 2/3 * 0.8 * 0.8 * 3/4 + 0.5 * 1/3 * 0.2 * 0.2 * 1/4 + 0.5 * 1/3 * 0.2 * 0.8 * 2/4`

```{r}
urn = c(0, 1) %>% 
  tibble() %>% 
  rename("urn" = ".")
```


```{r}
tr_0 = sample_n(urn, 1)
```


```{r}
addaball = function(df = urn, tr = tr_0, pra = 0.2, prb = 0.8) {
  if (pull(tr, urn) == 0) {
    outcome = rbinom(1, 1, pra)
    if (outcome == 0) {
      df = df %>% 
        add_row(urn = 1)
    }
    else {
      df = df %>% 
        add_row(urn = 0)
    }
  }
  if (pull(tr, urn) == 1) {
    outcome = rbinom(1, 1, prb)
    if (outcome == 0) {
      df = df %>% 
        add_row(urn = 0)
    }
    else {
      df = df %>% 
        add_row(urn = 1)
    }
  }
  return(df)
}
```

```{r, include = FALSE, eval = FALSE}
sim = function(func = addaball) {
  deck_df = c(0, 1) %>% 
    tibble() %>% 
    rename("urn" = ".")
  df = func() %>% 
    func() %>% 
    func()
  return(df)
}
```

```{r, include = FALSE, eval = FALSE}
urn1 = addaball()
urn1
```

```{r, include = FALSE, eval = FALSE}
tr_1 = sample_n(urn1, 1)
urn2 = addaball(df = urn1)
urn2
```

```{r}
pickaball = function(df = urn, func = addaball, tr, pra, prb) {
  tr_0 = sample_n(df, 1)
  urn1 = func(df, tr, pra = pra, prb = prb)
  tr_1 = sample_n(urn1, 1)
  urn2 = func(df = urn1, tr = tr_1, pra = pra, prb = prb)
  tr_2 = sample_n(urn2, 1)
  urn3 = func(df = urn2, tr = tr_2, pra = pra, prb = prb)
  tr_3 = sample_n(urn3, 1)
  res = rbind(tr_0, tr_1, tr_2, tr_3) %>% 
    rename("treatment" = "urn")
  return(res)
}
```

```{r ev}
sim = function(n = 10, df = urn, func = addaball, tr = tr_0, pra = 0.2, prb = 0.8) {
  set.seed(719)
  output = rerun(n, pickaball(df, func, tr, pra, prb))
  
  for (i in 1:length(output)) {
    output[[i]] = output[[i]] %>% 
      add_column("trial" = i)
  }
  
  nested_output = output %>% 
    bind_rows() %>% 
    nest(-trial)
  
  dist_df = nested_output %>% 
    mutate(data = map(data, sum))
  
  return(dist_df)
}
```

```{r, cache = TRUE}
dist_df = sim(n = 5000)
```


```{r}
dist_df %>% 
  unnest(cols = c(data)) %>% 
  pull(data) %>% 
  hist(right = FALSE, breaks = c(0:5), main = "Histogram of B", xlab = "B")
```

A simulation shows the distribution of outcomes of treatment B.

```{r}
exp_df = dist_df %>% 
  unnest(cols = c(data)) %>% 
  count(data) %>% 
  group_by(data) %>% 
  mutate(B = n/5000)
```

The expected value can be estimated as `r 0 * pull(exp_df, B)[[1]] + 1 * pull(exp_df, B)[[2]] + 2 * pull(exp_df, B)[[3]] + 3 * pull(exp_df, B)[[4]] + 4 * pull(exp_df, B)[[5]]`.

### Part b

```{r, cache = TRUE}
dist_df2 = sim(n = 5000, pra = 0.3, prb = 0.3)
```

```{r}
dist_df2 %>% 
  unnest(cols = c(data)) %>% 
  pull(data) %>% 
  hist(right = FALSE, breaks = c(0:5), main = "Histogram of B", xlab = "B")
```

```{r}
exp_df2 = dist_df2 %>% 
  unnest(cols = c(data)) %>% 
  count(data) %>% 
  group_by(data) %>% 
  mutate(B = n/5000)
```

The expected value can be estimated as `r 0 * pull(exp_df2, B)[[1]] + 1 * pull(exp_df2, B)[[2]] + 2 * pull(exp_df2, B)[[3]] + 3 * pull(exp_df2, B)[[4]] + 4 * pull(exp_df2, B)[[5]]`.