---
title: "20201005-p8133_probset2_jsg2145"
author: "Jared Garfinkel"
date: "10/5/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 2

## Part a

```{r}
14*pbinom(1, 14, .05) + 34*(1-pbinom(1, 14, .05))
```


The expected value of the sample size is 17.05 ~ 18.

## Part b

Go means that there was at least one response in the futility trial and 4 responses in the go-no-go trial.

```{r}
gonogo = function(stage1, stage2, null, alt, n1, n2) {
  res = vector(mode = "list", length = stage1)
  for(i in 1:stage1) {
    for(j in 1:stage2){
      res[[c(i,j)]] = bind_cols("n1_response" = i, "n2_response" = j-i)
    }
    res[[i]] = map(res[[i]], ~bind_rows(.))
  }

  sum_df = bind_rows(res) %>% 
    mutate(errorI = dbinom(n1_response, stage1, null) * dbinom(n2_response, stage2, null),
           power = dbinom(n1_response, stage1, alt) * dbinom(n2_response, stage2, alt)) %>%
    filter(n1_response > n1-1,
           n1_response + n2_response > n2-1) %>% 
    summarize(sum_error = sum(errorI),
              sum_power = sum(power))
  
  return(sum_df)
}
```

```{r}
sum_df = gonogo(stage1 = 14, stage2 = 20, null = .05, alt = 0.2, n1 = 1, n2 = 4)
```

The probability of rejecting the null given that the null is true is `r round(pull(sum_df, sum_error), 4)`.

## Part c

The probability of rejecting the null given that the alternative is true is `r round(pull(sum_df, sum_power), 4)`.

## Part d

type I error = .0804, power = 0.9041

null = 5%, alternative = 20%

n = $$\frac{2(z_{\alpha}~+~z_{1-\beta})^2(\sigma^2)}{\Delta^2}$$

```{r}
fixed = function(maxn, null, alt, error_prev, power_prev) {
  errorI = vector(mode = "list", length = maxn)
  power = vector(mode = "list", length = maxn)
  res = vector(mode = "list", length = maxn)
  for(i in 1:maxn) {
    for(j in 1:i) {
      errorI[[c(i, j)]] = 1 - pbinom(j-1, i, null)
      power[[c(i, j)]] = 1 - pbinom(j-1, i, alt)
      res[[c(i, j)]] = bind_cols("i" = i, "j" = j, "errorI" = errorI[[c(i, j)]], "power" = power[[c(i, j)]])
    }
    res[[i]] = map(res[[i]], ~bind_rows(.))
  }
  
  fixed_df = res %>% 
    bind_rows() %>% 
    filter(errorI < error_prev,
           power > power_prev)
  
  return(fixed_df)
}
```

```{r}
fixed(maxn = 100, null = 0.05, alt = 0.2, error_prev = 0.0804, power_prev = .9041)
```


The sample size 32 with a response of 4 or more subjects satisfies these parameters.

There is `r round(1 - pbinom(3, 32, .05), 5)` probability of rejecting the null with this outcome under the null.

Under the alternative, there is `r round(1 - pbinom(3, 32, .2), 4)` probability for rejecting this outcome.

The number of participants required for an adaptive design trial if the futility trial is successful is 2 more than for the adaptive model. 

However, the adaptive model has an expected sample size of only 17, which is 15 fewer than in the fixed design.