---
title: "20201105-p8133_probset8_jsg2145"
author: "Jared Garfinkel"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(dfcrm)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 8

Choose the prior to be the Gamma distribution. 

The exponential is $\lambda{e}^{-\lambda{x}}$.

So,

The posterior probability of $\lambda_1$ and $\lambda_2$ given $x_1,...,x_n$ and $y_1,...,y_m$ can be defined with the probability distribution function and the likelihood of x and y given $\lambda$.

$P(\lambda|x) = \frac{P(x|\lambda)}{P(x)}P(\lambda)$,

where P(x) = $$\int{p(x|\lambda)p(\lambda)d\lambda}$$

In other words: posterior probability $\propto$ likelihood x prior probability

The likelihood is $L(\lambda) = \prod{\lambda{exp(-\lambda}x)} = {\lambda}^ne^{-{\lambda}n\bar{x}}$

$p(\lambda) = \lambda^{n}exp(-{\lambda}n\bar{x})\frac{\beta^\alpha}{\Gamma(\alpha)}\lambda^{\alpha-1}exp(-\lambda\beta)$

Rearrange terms to get

= $\frac{\beta^\alpha}{\Gamma(\alpha)}\lambda^{(\alpha + n)-1}exp(-\lambda(\beta+n\bar{x}))$

The prior predictive distribution is the integral of the posterior distribution with respect to the prior.

So, the prior predictive distribution = $\frac{\beta^\alpha}{\Gamma(\alpha)}\lambda^{(\alpha + n)-1}exp(-\lambda(\beta+n\bar{x}))$

Removing the first term from the integral as it is not related to lambda, 

$$\frac{\beta^\alpha}{\Gamma(\alpha)}\int_\lambda{\lambda^{(\alpha + n)-1}exp(-\lambda(\beta+n\bar{x}))}$$

By multiplying through by a factor of $\frac{(\beta+n\bar{x})^{\alpha+n}}{\Gamma(\alpha+n)}$,

the integral can be written as:

$$\int_\lambda{\Gamma(\lambda; \alpha+n, \beta+n\bar{x})d\lambda}$$

This simplifies to 1 by the rules of exponential family distributions.

So,

p($\lambda$) = $$\frac{\beta^\alpha}{\Gamma(\alpha)}\frac{\Gamma(\alpha+n)}{(\beta+n\bar{x})^{\alpha+n}}$$

$$\frac{\beta^\alpha}{\Gamma(\alpha)}\frac{\Gamma(\alpha+n)}{(\beta+n\bar{x})^{\alpha+n}}$$

$$\left(\frac{\beta}{\beta+n\bar{x}}\right)^\alpha\left(\frac{1}{\beta+n\bar{x}}\right)^n\frac{\Gamma(\alpha+n)}{\Gamma(\alpha)}$$
```{r, include = FALSE, eval = FALSE}
So, the posterior probability of $\lambda_1$ is 

$$\frac{\beta^\alpha}{\Gamma(\alpha)}\lambda_1^{\alpha-1}e^{-\beta\lambda_1}{\lambda_1}e^{-\lambda_1{X_n}}$$

= $$\lambda_1^\alpha{e}^{-\lambda_1(\beta+X_n)}$$

and $\lambda_2$ will have a posterior probability of,

= $$\lambda_2^\alpha{e}^{-\lambda_2(\beta+Y_m)}$$

So, if $\frac{\lambda_1^\alpha{e}^{-\lambda_1(\beta+X_n)}}{\lambda_2^\alpha{e}^{-\lambda_2(\beta+Y_m)}}$ > 1, then the probability is true.

```