---
title: "20201107-p8133_probset11_jsg2145"
author: "Jared Garfinkel"
date: "11/7/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(dfcrm)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 11

## Part a

The first patient receives 1 of 2 potential treatments with equal probability.

p(A) = p(B) = 0.5

$p(tr = A|tr_{i-1} = A) = 0.2$

$p(tr = A|tr_{i-1} = B) = 0.8$

$p(tr = B|tr_{i-1} = A) = 0.2$

$p(tr = B|tr_{i-1} = B) = 0.8$

From this, the distribution of the first 4 patients can be calculated:

AAAA = 0.5 * 0.2 * 0.2 * 0.2 = 0.004
AAAB = 0.5 * 0.2 * 0.2 * 0.8 = 0.016
AABA = 0.5 * 0.2 * 0.8 * 0.2 = 0.016
ABAA = 0.5 * 0.8 * 0.2 * 0.2 = 0.016
BAAA = 0.5 * 0.2 * 0.2 * 0.2 = 0.004
AABB = 0.5 * 0.2 * 0.8 * 0.8 = 0.064
ABBA = 0.5 * 0.8 * 0.8 * 0.2 = 0.064
BBAA = 0.5 * 0.8 * 0.2 * 0.2 = 0.016
ABBB = 0.5 * 0.8 * 0.8 * 0.8 = 0.256
BBBA = 0.5 * 0.8 * 0.8 * 0.2 = 0.064
BBBB = 0.5 * 0.8 * 0.8 * 0.8 = 0.256
BBAB = 0.5 * 0.8 * 0.2 * 0.8 = 0.064
BABB = 0.5 * 0.2 * 0.8 * 0.8 = 0.064
BAAB = 0.5 * 0.2 * 0.2 * 0.8 = 0.016
BABA = 0.5 * 0.2 * 0.8 * 0.2 = 0.016
ABAB = 0.5 * 0.8 * 0.2 * 0.8 = 0.064

P(B = 0) = 0.004
P(B = 1) = 0.016 + 0.016 + 0.016 + 0.004 = .052
P(B = 2) = 0.064 + 0.064 + 0.016 + 0.016 + 0.016 + 0.064 = 0.24
P(B = 3) = 0.256 + 0.064 + 0.064 + 0.064 = .448
P(B = 4) = 0.256

0.004 + 0.052 + .24 + .256 + .448 = 1

```{r}
tibble(B = c(0:4),
       Pr = c(.004,0.052,0.24,0.448,0.256)) %>% 
  ggplot(aes(x = B, y = Pr)) +
  geom_bar(stat = "identity")
```

E[B] = 0 * 0.004 + 1 * 0.052 + 2 * 0.24 + 3 * 0.448 + 4 * 0.256 = `r 0 * 0.004 + 1 * 0.052 + 2 * 0.24 + 3 * 0.448 + 4 * 0.256`

A balanced design would have 2 patients in each treatment group compared to an average of 2.9 in Zelen's play-the-winner design.

## Part b

AAAA = 0.5 * 0.3 * 0.3 * 0.3 = 0.0135
AAAB = 0.5 * 0.3 * 0.3 * 0.7 = 0.0315
AABA = 0.5 * 0.3 * 0.7 * 0.7 = 0.0735
ABAA = 0.5 * 0.7 * 0.7 * 0.3 = 0.0735
BAAA = 0.5 * 0.7 * 0.3 * 0.3 = 0.0315
AABB = 0.5 * 0.3 * 0.7 * 0.3 = 0.0315
ABBA = 0.5 * 0.7 * 0.3 * 0.7 = 0.0735
BBAA = 0.5 * 0.3 * 0.7 * 0.3 = 0.0315
ABBB = 0.5 * 0.7 * 0.3 * 0.3 = 0.0315
BBBA = 0.5 * 0.3 * 0.3 * 0.7 = 0.0315
BBBB = 0.5 * 0.3 * 0.3 * 0.3 = 0.0135
BBAB = 0.5 * 0.3 * 0.7 * 0.7 = 0.0735
BABB = 0.5 * 0.7 * 0.7 * 0.3 = 0.0735
BAAB = 0.5 * 0.7 * 0.3 * 0.7 = 0.0735
BABA = 0.5 * 0.7 * 0.7 * 0.7 = 0.1715
ABAB = 0.5 * 0.7 * 0.7 * 0.7 = 0.1715

P(B = 0) = 0.0135
P(B = 1) = 0.0315 + 0.0735 + 0.0735 + 0.0315 = 0.21
P(B = 2) = 0.0315 + 0.0735 + 0.0315 + 0.0735 + 0.1715 + 0.1715 = .553
P(B = 3) = 0.0315 + 0.0315 + 0.0735 + 0.0735 = 0.21
P(B = 4) = 0.0135

0.0135 + 0.21 + 0.553 + 0.21 + 0.0135 = 1 

```{r}
tibble(B = c(0:4),
       Pr = c(0.0135,0.21,0.553,0.21,0.0135)) %>% 
  ggplot(aes(x = B, y = Pr)) +
  geom_bar(stat = "identity")
```

E[B] = 0 * 0.0135 + 1 * 0.21 + 2 * 0.553 + 3 * 0.21 + 4 * 0.0135 = `r 0 * 0.0135 + 1 * 0.21 + 2 * 0.553 + 3 * 0.21 + 4 * 0.0135`

When both treatments have a probability of response of 0.3 the expected value of the distribution of B is the same as the balanced design.