---
title: "p8133_probset_jsg2145"
author: "Jared Garfinkel"
date: "9/24/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(haven)
library(readxl)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 1

Given:

H_0 = 0
H_1 = -10
variance = 20


The power and type I error rate are needed to calculate the sample size.

We assume that the change in LDL is due to the medication and not some outside factor.

n = $$\frac{2(z_{\alpha}~+~z_{1-\beta})^2(\sigma^2)}{\Delta^2}$$

n = `r round((1.96+.8416)^2*20*2/(100), 4)`

The minimum sample size required is 4 in each group.

# Problem 2

## Part a

$$E[{responses}] = np$$ = 14 * .05 = 0.7

## Part b

Go means that there was at least one response in the futility trial and 4 responses in the go-no-go trial.

So, the probability of a go decision is the intersection of 1 response out of 14 and 3 responses out of 20.

The probability of 1 response in the futility trial is = `r round(1 - pbinom(1, 14, .05), 4)`.

The probability of 3 responses in the additional trial is = `r round(1 - pbinom(3, 20, .05), 4)`.

So the probability of a go result is = 0.153 * 0.0159 = `r round(.153*0.0159, 4)`.

## Part c

1-Pr($\ge 1 \cap \ge 3, p = 0.2$) = `r round(1 - pbinom(1, 14, .2)*(1-pbinom(3, 20, .2)), 4)`

## Part d

type I error = .0024, power = .8835

null = 5%, alternative = 20%

n = $$\frac{2(z_{\alpha}~+~z_{1-\beta})^2(\sigma^2)}{\Delta^2}$$

```{r}
errorI = vector(mode = "list", length = 100)
power = vector(mode = "list", length = 100)
res = vector(mode = "list", length = 100)
for(i in 1:100) {
  for(j in 1:i) {
    errorI[[c(i, j)]] = 1 - pbinom(j-1, i, 0.05)
    power[[c(i, j)]] = 1 - pbinom(j-1, i, 0.2)
    res[[c(i, j)]] = bind_cols("i" = i, "j" = j, "errorI" = errorI[[c(i, j)]], "power" = power[[c(i, j)]])
  }
  res[[i]] = map(res[[i]], ~bind_rows(.))
}

res %>% 
  bind_rows() %>%
  filter(errorI < 0.0024,
         power > 0.8835)
```

The sample size 67 with a response of 10 or more subjects satisfies these parameters.

There is `r round(1 - pbinom(10, 67, .05), 5)` chances of this outcome under the null.

Under the alternative, there is `r round(1 - pbinom(10, 67, .2), 4)` probability for this outcome.

